### 2016-11-18 Table1.tex
added table to repos

\documentclass[10pt]{article}
\usepackage{amsfonts}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\linespread{1}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}
\begin{document}

\begin{table}
\begin{tabular}{lccccccccc}
                                &  \multicolumn{9}{c}{Computer Software for Association Mapping} \\ \cline{2-10}
                                 & \multicolumn{6}{c}{Multiple-locus methods}  & & \multicolumn{2}{c}{Single-locus methods} \\  \cline{2-7}   \cline{9-10}
 Features                 & AMplus  & bigRR & glmnet & LMM-Lasso & MLMM & r2VIM     & & FaST-LMM & GEMMA \\  \hline
purpose built$^\dagger$    &   \cmark     &    \cmark      &  \xmark   &  \cmark  &  \cmark  &  \cmark && \cmark & \cmark           \\ [0.2cm]
 user  manual           & \cmark &   \xmark   & \cmark   & \xmark     &   \cmark     & \xmark          && \cmark    &    \cmark    \\  [0.2cm]
all snps fitted  &    \cmark     &    \cmark        &   \cmark          &      \cmark             &  \xmark          &  \xmark              &&   \xmark   &  \xmark     \\  
simultaneously      &         &            &             &                   &            &                &&      &      \\  [0.2cm]
capacity to include          &   \cmark   &      \cmark      &     \cmark    &  \xmark                 &      \cmark      &      \cmark          &&      \cmark & \cmark     \\  
other fixed effects$^\ddagger$                  &         &              &          &                    &            &                &&      &      \\  [0.2cm]
limited  memory$^\star$                   &   \cmark    &      \xmark      & \xmark          &  \xmark    &  \xmark    &   \xmark     && \cmark      & \xmark    \\  [0.2cm]
best set of snp           &    \cmark     &  \xmark    &  \xmark         &    \xmark               &     \cmark       &  \xmark              &&  \xmark   & \xmark      \\  
returned as result$^\uparrow$               &                           &        & & & & & & & \\  [0.2cm]
error checking           &  \cmark      &     \xmark        &      \xmark     &           \xmark        &      \xmark       &         \xmark       && \cmark  & \xmark \\ \hline

\end{tabular}
{$^\dagger$ \scriptsize{computer software specifically designed for the analysis of data from GWAS}}\\
{$^\ddagger$ \scriptsize{ allows  for the effect of population structure to be included as additional fixed effects}}\\
{\small $^\star$ \scriptsize{able to deal with data larger than the memory capacity of the computer}} \\
{\small $^\uparrow$ \scriptsize{no need to determine which 
marker loci are significant relative to some threshold or further analysis to calculate the significance of the findings. }}
\end{table}

\end{document}