### 2018-06-27 Table1_methodssummary.tex

\documentclass[10pt]{article}
\usepackage{amsfonts}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\linespread{1}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}
\begin{document}

\begin{table}
\caption{A summary of the features possessed by the Eagle package and the comparison implementations.}
\begin{tabular}{lccccccccc}
                                &  \multicolumn{9}{c}{Computer Software for Association Mapping} \\ \cline{2-10}
                                 & \multicolumn{6}{c}{Multiple-locus}  & & \multicolumn{2}{c}{Single-locus} \\  \cline{2-7}   \cline{9-10}
 Features                 & Eagle  & bigRR & glmnet & LMM-Lasso & MLMM & r2VIM     & & FaST-LMM & GEMMA \\  \hline
Purpose built$^1$    &   \cmark     &    \cmark      &  \xmark   &  \cmark  &  \cmark  &  \cmark && \cmark & \cmark           \\ [0.2cm]
Well documented$^2$           & \cmark &   \xmark   & \cmark   & \xmark     &   \cmark     & \xmark          && \cmark    &    \cmark    \\  [0.2cm]
Simultaneous   &    \cmark     &    \cmark        &   \cmark          &      \cmark             &  \xmark          &  \xmark              &&   \xmark   &  \xmark     \\  
fitting of SNPs     &         &            &             &                   &            &                &&      &      \\  [0.2cm]
Additional fixed effects$^3$          &   \cmark   &      \cmark      &     \cmark    &  \xmark                 &      \cmark      &      \cmark          &&      \cmark & \cmark     \\   [0.2cm]
Data larger than RAM$^4$                   &   \cmark    &      \xmark      & \xmark          &  \xmark    &  \xmark    &   \xmark     && \cmark      & \xmark    \\  [0.2cm]
Threshold free$^5$        &    \cmark     &  \xmark    &  \xmark         &    \xmark               &     \cmark       &  \xmark              &&  \xmark   & \xmark      \\  [0.2cm]
Informative error checking           &  \cmark      &     \xmark        &      \xmark     &           \xmark        &      \xmark       &         \xmark       && \cmark  & \xmark \\ \hline

\end{tabular}
{$^1$ \scriptsize{Computer software specifically designed for the analysis of data from GWAS.}}\\
{$^2$ \scriptsize{More than just a readme file or comments in an example file. Those programs with ticks had detailed user manuals.}}\\
{$^3$ \scriptsize{Ability to accommodate additional fixed effects in the model such as age, sex, and population structure effects.}} \\
{$^4$ \scriptsize{Able to deal with data larger than the memory capacity of the computer.}} \\
{$^5$ \scriptsize{Results reported as the set of SNP closest to the genes influencing a trait. No need to construct thresholds to determine 
significance of the findings.}}\\
{$^6$  \scriptsize{All the programs terminated on errors. However, not all the programs informed the user of the cause and how to fix the errors.}}\\
\end{table}

\end{document}